Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting

### 2020-01-08 pkg/circtree/inst/wind_paper/draft/draft.Rnw
Update abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2946 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: d409f23c44bb961f6411bdd6d1aa6471cccf40a2

\Abstract{
While circular data occur in a wide range of scientific fields, the methodology
for distributional modeling and probabilistic forecasting of circular response
variables is rather limited. Most of the existing methods are built on the
framework of generalized linear and additive models, which are often
challenging to optimize and to interpret. Therefore, we suggest circular
regression trees and random forests as an intuitive alternative approach that
is relatively easy to fit. Building on previous ideas for trees modeling
circular means, we suggest a distributional approach for both trees and forests
yielding probabilistic forecasts based on the von~Mises distribution. The
resulting tree-based models simplify the estimation process by using the
available covariates for partitioning the data into sufficiently homogeneous
subgroups so that a simple von~Mises distribution without further covariates
can be fitted to the circular response in each subgroup. These circular
regression trees are straightforward to interpret, can capture nonlinear
effects and interactions, and automatically select the relevant covariates that
are associated with either location and/or scale changes in the von~Mises
distribution. Combining an ensemble of circular regression trees to a circular
regression forest yields a local adaptive likelihood estimator for the
von~Mises distribution that can regularize and smooth the covariate effects.
The new methods are evaluated in a case study on probabilistic wind direction
forecasting at two Austrian airports, considering other common approaches as a
benchmark.
}

>>>>

\Abstract{
%While circular data occur in a wide range of scientific fields, the methodology
%for distributional modeling and probabilistic forecasting of circular response
%variables is rather limited. Most of the existing methods are built on the
%framework of generalized linear and additive models, which are often
%challenging to optimize and to interpret. Therefore, we suggest circular
%regression trees and random forests as an intuitive alternative approach that
%is relatively easy to fit. Building on previous ideas for trees modeling
%circular means, we suggest a distributional approach for both trees and forests
%yielding probabilistic forecasts based on the von~Mises distribution. The
%resulting tree-based models simplify the estimation process by using the
%available covariates for partitioning the data into sufficiently homogeneous
%subgroups so that a simple von~Mises distribution without further covariates
%can be fitted to the circular response in each subgroup. These circular
%regression trees are straightforward to interpret, can capture nonlinear
%effects and interactions, and automatically select the relevant covariates that
%are associated with either location and/or scale changes in the von~Mises
%distribution. Combining an ensemble of circular regression trees to a circular
%regression forest yields a local adaptive likelihood estimator for the
%von~Mises distribution that can regularize and smooth the covariate effects.
%The new methods are evaluated in a case study on probabilistic wind direction
%forecasting at two Austrian airports, considering other common approaches as a
%benchmark.
While circular data occur in a wide range of scientific fields, the methodology
for distributional modeling and probabilistic forecasting of circular response
variables is rather limited. Most of the existing methods are built on the
framework of generalized linear and additive models, which are often
challenging to optimize and interpret. Therefore, building on previous ideas
for trees modeling circular means, we suggest a distributional approach for
regression trees and random forests yielding probabilistic forecasts based on
the von~Mises distribution. The resulting tree-based models simplify the
estimation process by using the available covariates for partitioning the data
into sufficiently homogeneous subgroups so that a simple von~Mises distribution
without further covariates can be fitted to the circular response in each
subgroup. These circular regression trees are straightforward to interpret, can
capture nonlinear effects and interactions, and automatically select the
relevant covariates that are associated with either location and/or scale
changes in the von~Mises distribution. Combining an ensemble of circular
regression trees to a circular regression forest can regularize and smooth the
covariate effects. The new methods are evaluated in a case study on
probabilistic wind direction forecasting at two Austrian airports, considering
other common approaches as a benchmark.
}
